---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    theme: united
    keep_md: true
---
Ian F. McCallum
2014-10-19

```{r echo=FALSE}
library(knitr)
```

```{r setoptions echo=FALSE}
opts_chunk$set(fig.path="figure/")
```

**Prerequisites:** 
- The file 'activity.csv' is the working directory

### Loading and preprocessing the data

Read in the file 'activity.csv' and create a summary by day (called daily.activity), and a summary by invterval (called interval.activity)

```{r load.and.preprocess}
activity <- read.csv("activity.csv")

library(plyr)
library(ggplot2)

daily.activity <- ddply(activity, "date", summarise, daily.total=sum(steps))
interval.activity <- ddply(activity, "interval", 
                           summarise, mean.steps=mean(steps,na.rm=TRUE))

```

### Question 1: What is mean total number of steps taken per day?

**Prerequisites:**
- The plyr package has been loaded
- The ggplot2 package has been loaded
- The dataframe daily.activity has been created

To plot the histogram of the number of steps, ddply is used to aggregate 
the data for each day using the data frame

```{r daily.total.histogram}

p <- ggplot(daily.activity, aes(x=daily.activity$daily.total))
p <- p + geom_histogram(binwidth=2000, fill="orange", colour="lightgrey")
p <- p + ggtitle("Daily Step Total")
p <- p + xlab("Steps")

print(p)
```

## What is the average daily activity pattern?

**Prerequisites:**
- The plyr package has been loaded
- The ggplot2 package has been loaded
- The dataframe daily.activity has been created

To plot the activity by interveral, ddply is used to aggregate 
the data for each interval

```{r activity.patterns}

p <- ggplot(interval.activity, aes(x=interval, 
                                   y=mean.steps))
p <- p + geom_line()
p <- p + ggtitle("Interval Activity Patterns")
p <- p + xlab("Invterval")
p <- p + ylab("Steps")
print(p)

```

The interval with the most activity is ```{r}

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
